An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-05-06 EconGrowth.Rnw
Changes to to-do list and other small adjustments.

\item{Heun: Include the adjusted sankey diagram in the seminar}
\item{Heun: Add the insight that for indexed factors of production $q$ and $x$, there is small difference in absolute magnitude and even smaller difference in indexed values unless $f_{exergy-allocation}$ is significantly different from 1.0 \emph{and} there is a shift in energy input to the economy.}

@@

@@

\begin{itemize}
  \item{The use of chemical exergy, as opposed to thermal energy, in economic growth models is probably unnecessary. Table \ref{tab:Exergy_Allocation} shows that exergy allocation factors are all close to unity. In the absence of very large changes in energy mix for an economy, indexed chemical exergy ($x$) and thermal energy ($q$) are very similar, and Figure \ref{fig:Historical_Lattice_Graph} shows that $q$ and $x$ lines lie atop one another for almost every economy.}
  \item{Figure \ref{fig:CDeGridLattice} shows that the set of parameters ($\lambda$, $\alpha$, $\beta$, and $\gamma$) that minimizes $SSE$ for the energy-augmented Cobb-Douglas production function always (with the exception of Saudi Arabia) has $\beta = 0$ or $\gamma = 0$. This result implies that it is unnecessary to include both labor ($l$) and energy ($e$) in the Cobb-Douglas production function.}
  \item{Partial residual plots for both the Cobb-Douglas model (Figures \ref{fig:CDPartialResidualsQ}--\ref{fig:CDPartialResidualsU} and CES model (Figures \ref{fig:CESPartialResidualsQ}--\ref{fig:CESPartialResidualsU}) indicate that augmenting the Cobb-Douglas and CES production functions with energy is not beneficial.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} shows that important energy parameters for both the energy-augmented Cobb-Douglas and CES production functions ($\gamma$ and $\delta$, respectively) have large confidence intervals such that we are unable to say (at the 95\% confidence level) that energy is an important factor for economic growth, vis-\`{a}-vis capital stock ($k$) or labor ($l$) for most economies.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} indicates that energy (in all quantifications: $q$, $x$, and $u$) is complementary to capital stock ($k$) and labor ($l$) for producing economic growth, according to the CES model.}
  \item{Figures \ref{fig:LINEXq_Params_Graph}--\ref{fig:LINEXu_Params_Graph} indicate that LINEX model fitting parameters $a_0$ and $c_t$ are significant (at the 95\% confidence level) for nearly all countries. Figure \ref{fig:LINEX_GDP_Lattice_Graph} shows that the LINEX production function provides reasonably good fits to historical data. This evidence indicates that the LINEX production function (with energy ($e$) as its only factor of production) is a reasonable replacement the Cobb-Douglas and CES production functions which have capital stock ($k$) and labor ($l$) as factors of production.}
  \item{****** Discuss the AIC values for each of the model types. It could happen that the LINEX models have a higher AIC value than the CD or CES models. If so, that would mean that the LINEX model is actually preferred to the CD or CES models, primarily because it does a good-enough job of predicting economic output with fewer adjustable parameters (2) compared to CD (2), CDe (3), CES (4), CESe (6). Discuss with Randy Pruim during Summer 2013.}
  \item{The concept of useful work ($u$) has potential and should be explored beyond developed economies. Figure \ref{fig:Historical_Lattice_Graph} shows that $u$ grows at very different rates from $k$ or $l$ for the developed economies. The covariance values in Tables \ref{tab:Covariance_US}--\ref{tab:Covariance_ZM} show that $u$ has much lower covariances with $k$ and $l$ than $q$ or $x$ has with $k$ and $l$. Thus, $u$ may contain more information relative to $k$ and $l$ than $q$ or $x$ and may, therefore, be more useful in economic growth modeling than $q$ or $x$. ******* Check this assumption after we obtain covariance values for $u$.**********}
\end{itemize}

>>>>

\item{Heun: Include Caleb's adjusted sankey diagram in the seminar}

@@

\item{\st{Heun: Add the insight that for indexed factors of production $q$ and $x$, there is small difference in absolute magnitude and even smaller difference in indexed values unless $f_{exergy-allocation}$ is significantly different from 1.0 \emph{and} there is a shift in energy input to the economy.}}

@@

\begin{itemize}
  \item{The use of chemical exergy, as opposed to thermal energy, in economic growth models is probably unnecessary. Table \ref{tab:Exergy_Allocation} shows that exergy allocation factors are all close to unity. Therefore, there is a small difference in absolute magnitude between $Q$ and $X$. However, there is no difference in indexed magnitude between $q$ and $x$, unless there are large changes in energy mix for an economy. Figure \ref{fig:Historical_Lattice_Graph} shows that $q$ and $x$ lines lie atop one another for almost every economy.}
  \item{The concept of useful work ($u$) is potentially useful and should be explored beyond developed economies. Figure \ref{fig:Historical_Lattice_Graph} shows that $u$ grows at very different rates from thermal energy ($q$) or exergy ($x$). Useful work may contain different information from thermal energy and exergy.}
  \item{Figure \ref{fig:CDeGridLattice} shows that the set of parameters ($\lambda$, $\alpha$, $\beta$, and $\gamma$) that minimizes $SSE$ for the energy-augmented Cobb-Douglas production function always (with the exception of Saudi Arabia) has $\beta = 0$ or $\gamma = 0$. This result implies that it is unnecessary to include both labor ($l$) and energy ($e$) in the Cobb-Douglas production function.}
  \item{Partial residual plots for both the Cobb-Douglas model (Figures \ref{fig:CDPartialResidualsQ}--\ref{fig:CDPartialResidualsU} and CES model (Figures \ref{fig:CESPartialResidualsQ}--\ref{fig:CESPartialResidualsU}) indicate that augmenting the Cobb-Douglas and CES production functions with energy is not beneficial.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} shows that important energy parameters for both the energy-augmented Cobb-Douglas and CES production functions ($\gamma$ and $\delta$, respectively) have large confidence intervals such that we are unable to say (at the 95\% confidence level) that energy is an important factor for economic growth, vis-\`{a}-vis capital stock ($k$) or labor ($l$) for most economies.}
  \item{Figure \ref{fig:CIvsParam_Lattice_Graph} indicates that energy (in all quantifications: $q$, $x$, and $u$) is complementary to capital stock ($k$) and labor ($l$) for producing economic growth, according to the CES model.}
  \item{Figures \ref{fig:LINEXq_Params_Graph}--\ref{fig:LINEXu_Params_Graph} indicate that LINEX model fitting parameters $a_0$ and $c_t$ are significant (at the 95\% confidence level) for nearly all countries. Figure \ref{fig:LINEX_GDP_Lattice_Graph} shows that the LINEX production function provides reasonably good fits to historical data. This evidence indicates that the LINEX production function (with energy ($e$) as its only factor of production) is a reasonable replacement the Cobb-Douglas and CES production functions which have capital stock ($k$) and labor ($l$) as factors of production.}
  \item{****** Discuss the AIC values for each of the model types. It could happen that the LINEX models have a higher AIC value than the CD or CES models. If so, that would mean that the LINEX model is actually preferred to the CD or CES models, primarily because it does a good-enough job of predicting economic output with fewer adjustable parameters (2) compared to CD (2), CDe (3), CES (4), CESe (6). Discuss with Randy Pruim during Summer 2013.}
\end{itemize}
